---
output: html_document
---

#### **To: Manager Chris Lee**  
  
#### **From: Wei Li & Florence Yan**  
  
#### **Date: `r format(Sys.time(), '%B %d, %Y')`**   
  
#### **Subject: Office product profitability analysis**  

***

```{r echo = FALSE}
# Course: BUAN 5210
# Title: Profitability analysis
# Purpose: To increase overall profit of the whole company
# Date: Jan 30, 2019
# Author: Wei Li & Florence Yan

```


```{r echo = FALSE}
# Clear environment of variables and functions
rm(list = ls(all = TRUE)) 

# Clear environmet of packages
if(is.null(sessionInfo()$otherPkgs) == FALSE)lapply(paste("package:", names(sessionInfo()$otherPkgs), sep=""), detach, character.only = TRUE, unload = TRUE)

```


```{r echo = FALSE, message = FALSE, warning = FALSE}
# Load librarys
library(tidyverse)
library(ggplot2)
```


```{r echo = FALSE}
# Load saved image of the office Tech Appendix workspace
load(".RData")
```


Our company has achieved good success since 2012. We have built good reputations and trust among customers. We feel proud of our work and company. However, we are now facing increasingly fierce competitions from online retailers, such as Amazon, Ebay or other platforms. What we can do is to keep our advantages and improve the aspects where we are not good at. Thus, it is worthwhile to analyze the historical sales data to find out possible monetization opportunities and identify potential drawbacks. In this study, our team explored several possible ways to improve profitability and the unseen problems we might have. 

After diving into the data, we mainly felt interested in two questions: 

+ What are the most profit-potential subcategories in the for regions?
+ How does discount affect profits? 

Firstly, we find out four subcategories that might interest our regional managers most, which are Tables, Furnishings, Copiers, and Binders, as the profits made in the four subcategories have the widest range across four regions. Since the four regions are geographically close to each other, they share similar characteristics in market potentials and thus can be comparable. In this case, if the profit of one subcategory is significantly different with other regions, the highest-profiting region may share the sales strategies with others. Thus the overall profit of the whole company might be able to improve. According the figure below, which shows the profitability of the four regions in four subcategories, our suggestions are as follows. First, south, east, and central regions need to learn how to improve profits of tables. In fact, based on our analysis, we suggest to reduce the sales of table since 3/4 of the regions are losing money on it. Secondly, central region needs to find out the reasons for the negative profit in the Furnishings and Binders sales. Thirdly, south region might need to learn how to make Copiers more profitable by communicating with other three regions. Fourthly, managers from west and east regions can share their sales strategies with others since they achieved good profits according o the data.

```{r echo = FALSE, include = TRUE, message = FALSE}
range_subCat <- max_min %>% 
  mutate(diff = max_prof - min_prof) %>% 
  arrange(desc(diff)) %>% 
  top_n(4)

subCat_to_see <- range_subCat$Sub.Category # 4 categories of biggest interest to managers

sale %>% 
  group_by(Sub.Category, Region) %>% 
  summarise(prof_sum = sum(Profit)) %>% 
  filter(Sub.Category %in% subCat_to_see) %>% 
  ggplot(aes(Sub.Category, prof_sum, fill = Region))+
  geom_bar(stat = 'identity',
           position = 'dodge') +
  coord_flip() +
  ylab('Total profit in the 4 subcategories') +
  xlab('4 subcategories that most interest managers') +
  ggtitle("Categories with widest profit range",
          sub = "The lower-profit area might learn from the highest-profit area about the sale \nstrategies and methods.") 

```

Our second finding shows that over half of our sales across four regions have discounts. However, regions with high profits like West doesn't offer as many discounts as low-profit regions. Thus, we think the latters should adjust their discount strategies since more discounts appearently didn't motivate higher profits. The figure below shows that the sales of Technology in central region which offered the highest discount as eighty percent indeed lost the most money across four regions.We recommand that company should find other ways like reducing the sales of unpopular categories and focusing on products with good performances to achieve higher profits.

```{r echo = FALSE, include = TRUE}
sale %>%
  ggplot(aes(x=Discount, Profit, color = Region)) +
  geom_point() + ggtitle("Profit vs. Discount", 
                         sub = "The Profit trend is going down as Piscount increases.")
```

In conclusion, we suggest managers from different regions to communicate and share with each other about their marketing strategies. Moreover, four subcategories includes tables, furnishings, copier, binders, and applications can be marked as target products since they did really well in some regions while others lost money on them. Secondly, we recomend that regions with low profits should adjust their discount strategies as they lost money with giving very high discounts. Please do not hesitate to contact me if you have questions or would like to discuss the analysis further. I can also provide my complete analysis with data and code if you would like more detail.


Best regards,

Florence Yan & Wei Li
